### 2024-01-23 outputs/paper/paper.qmd
Update paper.qmd

---
title: "Exploring Respiratory Infection Trends"
subtitle: "A Data Analysis of Post-COVID-19 Impacts in Toronto (2020-2023)"
author: 
  - Hailey Jang
thanks: "Code and data are available at: https://open.toronto.ca/dataset/outbreaks-in-toronto-healthcare-institutions/."
date: January 23, 2024
date-format: long
abstract: "This research delves into the trajectory of respiratory infections in Toronto in the aftermath of COVID-19, utilizing datasets from 2020 and 2023. Simulated data on respiratory infection rates reveal a discernible surge in respiratory ailments post-pandemic. Statistical examinations are applied to compare infection rates, providing valuable insights into the changing dynamics of respiratory health. The study contributes to a holistic understanding of the enduring repercussions of COVID-19 on respiratory well-being within the Toronto population."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction
The aftermath of the COVID-19 pandemic prompts an in-depth examination of its enduring impact, with a particular focus on respiratory health. This study centers its exploration on Toronto, a diverse urban hub, utilizing datasets from pivotal years—2020 and 2023. Simulated data on respiratory infection rates form the basis of our analysis, uncovering discernible patterns in the prevalence of respiratory ailments post-pandemic. Through statistical tests, we aim to elucidate the magnitude of these changes, contributing to a nuanced understanding of the evolving landscape of respiratory health in Toronto.

This research not only addresses local implications but also contributes valuable insights to the broader discourse on post-COVID-19 health dynamics. By unravelling the complexities of respiratory infections in Toronto, we strive to offer actionable insights for public health strategies, thereby mitigating the lasting impact of the pandemic on respiratory well-being.

# Data {#sec-data}
All the data utilized in this paper were sourced from the Toronto Open Data portal [@torontodata]. Specifically, The dataset is titled 'Outbreaks in Toronto Healthcare Institutions'. The cleaning and analysis processes were executed through the use of R, a statistical programming software [@citeR], along with several libraries including `tidyverse` [@tidy], `ggplot2` [@ggplot], `dplyr` [@dp], `readr` [@read], `tibble` [@tib], `janitor` [@jan], `reshape2` [@reshape], `knitr` [@knit], `ggbeeswarm` [@gg], `ggrepel` [@repel], and `here` [@here].