### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

---
title: "Unveiling Toronto's Respiratory Landscape"
subtitle: "Analyzing Post-COVID-19 Infection Trajectory and Rhinovirus Surges through 2020 and 2023 Datasets"
author: 
  - Hailey Jang
thanks: "Code and data supporting this analysis are available at: https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git"
date: January 25, 2024
date-format: long
abstract: "This research investigates the impact of COVID-19 on respiratory infections in Toronto, utilizing datasets from 2020 and 2023. Through analysis of simulated data on respiratory infection rates, with a focus on Rhinovirus, a noticeable surge in respiratory ailments post-pandemic is revealed. Statistical examinations comparing infection rates offer insights into the changing dynamics of respiratory health in the city. This study contributes valuable information towards understanding the lasting effects of COVID-19 on respiratory well-being in the Toronto population, providing a crucial foundation for public health strategies and interventions."
output: pdf_document
number-sections: true
format: pdf
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
```

# Introduction

In response to the global landscape following the COVID-19 pandemic, this study thoroughly explores its impact on public health, specifically concentrating on respiratory infections in Toronto. Through a meticulous examination of simulated data spanning 2020 to 2023, the research provides a nuanced understanding of the trajectory post the initial pandemic wave, with a specific focus on the prevalence of Rhinovirus. The results highlight a significant increase in respiratory ailments, with Rhinovirus emerging as a notable contributing factor.

Amidst global efforts to address the aftermath of the transformative COVID-19 pandemic, a noticeable gap exists in understanding long-term consequences, especially concerning respiratory health trends in urban populations. This paper aims to fill this gap by presenting a comprehensive analysis based on simulated data, scrutinizing respiratory infection rates in Toronto. The data section meticulously outlines data collection and analysis procedures, while the results section succinctly articulates key findings. The ensuing discussion contextualizes these findings within the broader framework of public health, shedding light on the evolving dynamics of respiratory health in a post-COVID-19 era.

As societies globally grapple with the intricate challenges posed by the aftermath of COVID-19, this study addresses the critical need for insights into enduring impacts on respiratory health, particularly in urban settings. Beyond empirical contributions, the study strives to play a proactive role in shaping future policies and basic infection control in real-life situations, serving as a guide for fortifying community health in the wake of unprecedented global events.

# Data 
All the data used in this study were sourced from the Toronto Open Data portal [@opendatatoronto]. The dataset, titled 'Outbreaks in Toronto Healthcare Institutions' serves as the foundation for our analyses.

For the purpose of data analysis, simulation, cleansing, and visualization, the statistical programming language R was judiciously employed [@citeR]. Additionally, to facilitate these analytical processes, a suite of specialized R packages was utilized, including but not limited to here [@citeggplot2] and the comprehensive tidyverse suite [@tidyverse].

## Respiratory Infections in 2020
Examining reports on respiratory outbreaks in Toronto healthcare institutions from 2020 to 2023 unveils intriguing patterns that provide valuable insights into the dynamic nature of infectious diseases. In Figure 1, which represents the year 2020, a consistent trend emerges, showcasing that respiratory infections consistently surpass Enteric illnesses. Out of 553 reported cases, 519 are respiratory, constituting approximately 93.85% of the total cases. Enteric outbreaks rank as the second most common with 33 cases, making up about 5.97%. There is only 1 case categorized as Other, representing 0.18%, and no cases are listed under Others. This underscores the persistent challenges posed by respiratory health issues, with respiratory outbreaks significantly outnumbering other types. Considering the early onset of the COVID-19 outbreak, the prevalence of respiratory outbreaks in this year aligns with historical expectations.

```{r}
#| echo: false
#| eval: true
#| label: fig-1
#| fig-cap: "Reports of Respiratory Outbreaks (2020)"
#| warning: false

library(tidyverse)

# Read the data
url <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/2bbbb602-f621-4f78-af5c-0d604fd63831"
respiratory_data_2020 <- read_csv(url, show_col_types = FALSE,
                                  col_types = cols(
                                    `Date Outbreak Began` = col_date(),
                                    `Date Declared Over` = col_date()
                                  ))

# Select relevant columns
selected_columns <- c("Institution Name", "Type of Outbreak", "Causative Agent - 1", "Date Outbreak Began", "Date Declared Over")
filtered_data <- respiratory_data_2020 %>% select(all_of(selected_columns))

# Plotting
ggplot(filtered_data, aes(x = `Type of Outbreak`, fill = `Type of Outbreak`)) +
  geom_bar() +
  labs(title = "Reports of Respiratory Outbreaks (2020)",
       x = "Type of Outbreak",
       y = "Count") +
  theme_minimal()

```

## Respiratory Infections in 2023
In contrast, Figure 2, representing the post-COVID period in 2023, unfolds a more complex narrative. While the underlying trend of respiratory infections surpassing Enteric illnesses persists, there is a significant increase in the frequency of respiratory outbreaks compared to the previous year. In the dataset, respiratory cases total 1016, with only three cases listed under 'Others'. There are also 47 cases of Enteric illnesses, signifying that almost 100% of the outbreaks are of a respiratory nature. This notable surge in respiratory infection rates may be attributed to various factors, including the lingering effects of the COVID-19 pandemic. The classification of 2023 as post-COVID implies a potential connection between the ongoing impact of the pandemic and the heightened prevalence of respiratory infections.

```{r}
#| echo: false
#| eval: true
#| label: fig-2
#| fig-cap: "Reports of Respiratory Outbreaks (2023)"
#| warning: false

library(tidyverse)

# Read the data
url_2023 <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/15ac28a1-ece5-4a97-8e72-11227be9f4f7"
respiratory_data_2023 <- read_csv(url_2023, show_col_types = FALSE,
                                  col_types = cols(
                                    `Date Outbreak Began` = col_date(),
                                    `Date Declared Over` = col_date()
                                  ))

# Select relevant columns
selected_columns_2023 <- c("Institution Name", "Type of Outbreak", "Causative Agent-1", "Date Outbreak Began", "Date Declared Over")
filtered_data_2023 <- respiratory_data_2023 %>% select(all_of(selected_columns_2023))

# Plotting
ggplot(filtered_data_2023, aes(x = `Type of Outbreak`, fill = `Type of Outbreak`)) +
  geom_bar() +
  labs(title = "Reports of Respiratory Outbreaks in Toronto Healthcare Institutions (2023)",
       x = "Type of Outbreak",
       y = "Count") +
  theme_minimal()
```
## Prevalence of Respiratory Infections
The analysis of respiratory infection trends in the aftermath of the COVID-19 pandemic reveals a noticeable increase in the overall prevalence of respiratory infections throughout the year 2023, presenting a distinct contrast with the preceding year, 2020. This notable rise extends beyond pre-pandemic norms, signifying a substantial transformation in the landscape of respiratory health, as highlighted in (@Maison). Consequently, there is a compelling need for an in-depth exploration of potential contributing factors and the development of adaptive public health strategies to effectively address the evolving challenges within the realm of respiratory infections.

# Results 
## Rhinovirus
In delving deeper into the landscape of respiratory outbreaks in Toronto healthcare institutions, a crucial aspect of our analysis involves identifying the specific viruses that contribute most significantly to these incidents. Understanding the predominant viral agents responsible for respiratory infections is pivotal for targeted preventive measures, effective treatment strategies, and informed public health interventions. In this exploration, we aim to unravel the intricate network of viral contributors, shedding light on the specific viruses that exert a substantial impact on the respiratory health dynamics within these healthcare settings. By discerning the key viral players, we can glean insights that will not only enhance our comprehension of the current state but also inform future strategies for mitigating the prevalence and consequences of respiratory outbreaks.

The graphical representation, @fig-3, comparing the rates of causative agents for respiratory outbreaks in Toronto healthcare institutions in 2020 and 2023 provides insightful observations. Notably, the data underscores a significant shift in the prominence of Rhinovirus as a causative agent between the two years. In 2023, Rhinovirus emerges as the predominant viral contributor, reporting a notably higher incidence compared to its occurrence in 2020.

Specifically, in 2023, Rhinovirus stands out as the most frequently reported causative agent for respiratory outbreaks, with 77 documented cases. This substantial increase is particularly noteworthy when contrasted with its presence in 2020, where Rhinovirus was reported merely seven times. The surge in Rhinovirus cases in 2023 positions it at the forefront, signifying a noteworthy change in the landscape of respiratory infections during this period.

Comparatively, in 2020, Respiratory syncytial virus took precedence as the leading causative agent, reporting 19 cases, followed closely by Influenza A with 18 cases. However, the dynamic shift in 2023, with Rhinovirus surpassing these previously dominant agents, suggests an evolving pattern in the prevalence of respiratory viruses. The observed rise in Rhinovirus cases may be indicative of changing epidemiological factors, altered patient demographics, or other environmental influences that warrant further investigation.

```{r}
#| echo: false
#| eval: true
#| label: fig-3
#| fig-cap: "Rates of Causative Agents (2020 vs. 2023)"
#| warning: false

library(tidyverse)

# Function to read and clean data
read_and_clean_data <- function(url, causative_agent_column) {
  data <- read_csv(url)

  # Exclude COVID-19 and Coronavirus from causative agents
  cleaned_data <- data %>%
    filter(!grepl("COVID-19|Coronavirus", data[[causative_agent_column]], ignore.case = TRUE))

  return(cleaned_data)
}

# Read and clean data for 2020
url_2020 <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/2bbbb602-f621-4f78-af5c-0d604fd63831"
cleaned_data_2020 <- read_and_clean_data(url_2020, "Causative Agent - 1")

# Read and clean data for 2023
url_2023 <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/15ac28a1-ece5-4a97-8e72-11227be9f4f7"
cleaned_data_2023 <- read_and_clean_data(url_2023, "Causative Agent-1")

# Standardize column names
cleaned_data_2020 <- cleaned_data_2020 %>%
  rename(Causative_Agent = "Causative Agent - 1")

cleaned_data_2023 <- cleaned_data_2023 %>%
  rename(Causative_Agent = "Causative Agent-1")

# Combine the datasets
combined_data <- bind_rows(
  mutate(cleaned_data_2020, Year = "2020"),
  mutate(cleaned_data_2023, Year = "2023")
)

# Plotting
ggplot(combined_data, aes(x = Causative_Agent, fill = Year)) +
  geom_bar(position = "dodge") +
  labs(x = "Causative Agent",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The @fig-3 highlights a substantial increase in Rhinovirus cases in 2023, positioning it as the primary causative agent for respiratory outbreaks in Toronto healthcare institutions. This observation prompts further exploration into the factors contributing to this shift and underscores the importance of ongoing surveillance and research to adapt public health strategies to the evolving landscape of respiratory infections.

# Discussion
## Rhinovirus in 2023
The resurgence of Rhinovirus (RV) in 2023, particularly in comparison to other respiratory viruses following the relaxation of social distancing measures, demands a detailed examination. At the forefront of this resurgence is the unique protein capsid of Rhinovirus, endowing it with heightened resistance to ethanol-containing disinfectants—an attribute not shared by viruses with lipid envelopes.The efficacy of facemasks in filtering out Rhinovirus is notably compromised due to its smaller size compared to influenza virus and coronavirus, potentially contributing to the observed resurgence. The implementation of social-distancing measures, restrictions on social gatherings, and intensified hygiene protocols in response to the March 2020 COVID-19 outbreak significantly curtailed exposure opportunities to Rhinovirus and other infectious agents. This led to a substantial reduction in both respiratory viral infections and healthcare utilization for asthma. However, the intricacies of this dynamic become apparent with the specific resurgence of Rhinovirus infections following the fall 2020 reopening of schools. Unlike other respiratory viruses that remained subdued, Rhinovirus exhibited a significant uptick in prevalence.

The discernible influence of social interventions on Rhinovirus underscores the significance of its intrinsic attributes, particularly its diminutive size and capsid coat. According to Kreger (@Kreger), these characteristics may contribute to its reduced susceptibility to masking agents and disinfectants. The observed incremental surge in asthma-related hospitalizations subsequent to the reopening of schools implies a complex interconnection between Rhinovirus infections and exacerbations of asthma. Significantly, notwithstanding the increase in admissions, the levels remained below those observed prior to the onset of the COVID-19 pandemic, revealing a nuanced interplay among respiratory infections, societal dynamics, and health outcomes associated with asthma.

\newpage

# Conclusion
The resurgence of Rhinovirus post-social distancing measures unveils a complex interplay of factors, highlighting the unique characteristics of Rhinovirus, the absence of influenza infections, and the dynamics of viral interference. The differential impact of social measures on various respiratory viruses underscores the necessity for tailored public health strategies. The observed partial increase in asthma hospitalizations following school reopening emphasizes the intricate relationship between respiratory infections and asthma outcomes.

Moving forward, a nuanced understanding of these dynamics is essential for mitigating the impact of respiratory infections and optimizing public health interventions in the post-COVID era. This exploration not only contributes to our comprehension of Rhinovirus resurgence but also underscores the need for adaptive strategies in the ongoing battle against respiratory infections in healthcare settings.

# Reference